Fast, Scalable Phrase-Based SMT Decoding

### 2016-02-09 acl.2016/acl2016.tex
acl paper

We trained a phrase-based system using the Moses toolkit with standard settings. The training data consisted of most of the publicly available Arabic-English data from Opus~\cite{tiedemann2012parallel} containing over 69 million parallel sentences, and tuned on a held out set. The phrase-table was then pruned, keeping only the top 100 entries per source phrase, according to $p(t|s)$. All models files were then binarized; the language models were binaized using KenLM~\cite{Heafield-kenlm}, the phrase table using the Probing phrase-table, lexicalized reordering model using the compact datastructure. These binary formats were choosen for their best-in-class multithreaded performance. Table~\ref{tab:model-files} gives details of the resultant sizes of the model files. For verification with a different dataset, we also occasionally used a second system trained on the French-English Europarl corpus (2m parallel sentences). 
\begin{table}[h]
\begin{center}
\begin{tabular}{|l|c|c|} \hline
		& ar-en	& fr-en \\ \hline
Phrase table  	& 17 	& 5.8 \\
Language model 	& 3.1  	& 1.8 \\ 
Lex-re model	& 2.3	& 637MB \\ \hline
\end{tabular}
\end{center}
\caption{Model sizes in GB}
\label{tab:model-files}
\end{table}
For testing decoding speed, we used a subset of the training data, Table~\ref{tab:test-sets}. The two test scenarios have differing characterics that we are interested in analyzing, ar-en have short sentences with large models while fr-en have overly long sentences with smaller models. Where we need to compare model scores, we used held out test sets.

@@

We create a dynamic memory pool which can grow as more memory is requested. The memory is not released, instead the pool can be reset in order for the memory to be re-used. We instantiate two pools for each thread, one which is never reset and another which is reset after the decoding each sentence. Objects are created in either pool according to their life cycle.

For critical objects with high churn such as hypotheses, thread-specific LIFO queues are used to recycle objects which are no longer used. This not only reduces memory wastage but re-uses recent objects which are likely to be in the CPU cache memory.

@@

\begin{figure}[h]
\centering
\begin{tabular}{cc}
{\includegraphics[scale=0.4]{scalability.png}} 
\end{tabular}
\caption{Decoding speed}
\label{fig:speed}
\end{figure}

@@

This contrast with Moses where speed increases to approximately 16 threads but then actually become slower if more threads are used, Figure~\ref{fig:moses-speed}.
\begin{figure}[h]
\centering
\begin{tabular}{cc}
{\includegraphics[scale=0.4]{moses-scalability.png}} 
\end{tabular}
\caption{Moses decoding speed}
\label{fig:moses-speed}
\end{figure}

>>>>

We trained a phrase-based system using the Moses toolkit with standard settings. The training data consisted of most of the publicly available Arabic-English data from Opus~\cite{tiedemann2012parallel} containing over 69 million parallel sentences, and tuned on a held out set. The phrase-table was then pruned, keeping only the top 100 entries per source phrase, according to $p(t|s)$. All models files were then binarized; the language models were binaized using KenLM~\cite{Heafield-kenlm}, the phrase table using the Probing phrase-table, lexicalized reordering model using the compact datastructure. These binary formats were choosen for their best-in-class multithreaded performance. Table~\ref{tab:model-files} gives details of the resultant sizes of the model files. For verification with a different dataset, we also occasionally used a second system trained on the French-English Europarl corpus (2m parallel sentences). 
\begin{table}[h]
\begin{center}
\begin{tabular}{|l|c|c|} \hline
		& ar-en	& fr-en \\ \hline
Phrase table  	& 17 	& 5.8 \\
Language model 	& 3.1  	& 1.8 \\ 
Lex re. model	& 2.3	& 637MB \\ \hline
\end{tabular}
\end{center}
\caption{Model sizes in GB}
\label{tab:model-files}
\end{table}
For testing decoding speed, we used a subset of the training data, Table~\ref{tab:test-sets}. The two test scenarios have differing characterics that we are interested in analyzing, ar-en have short sentences with large models while fr-en have overly long sentences with smaller models. Where we need to compare model scores, we used held out test sets.

@@

We create a dynamic memory pool which can grow as more memory is requested. The memory is not released, instead the pool can be reset in order for the memory to be re-used. We instantiate two pools for each thread, one which is never reset and another which is reset after the decoding each sentence. Datastructures are created in either pool according to their life cycle.

For critical datastructures with high churn such as hypotheses, thread-specific LIFO queues are used to recycle objects which are no longer used. This not only reduces memory wastage but re-uses recent objects which are likely to be in the CPU cache memory.

@@

\begin{figure}[h]
\centering
\begin{tabular}{cc}
{\includegraphics[scale=0.4]{scalability.png}} 
\end{tabular}
\caption{Our decoder's decoding speed}
\label{fig:speed}
\end{figure}

@@

This contrast with Moses where speed increases to approximately 16 threads but then actually become slower if more threads are used, Figure~\ref{fig:moses-speed}.
\begin{figure}[h]
\centering
\begin{tabular}{cc}
{\includegraphics[scale=0.4]{moses-scalability.png}} 
\end{tabular}
\caption{Moses' decoding speed}
\label{fig:moses-speed}
\end{figure}
